### 2021-05-06 written/eLifeRevision/Bittner_eLife2020.tex
latest revision

@@

Equipped with this method, we demonstrate the capabilities of EPI and present novel theoretical findings from its analysis.
\q{intro_results1}{First, we show EPI's ability to handle biologically realistic circuit models using a five-neuron model of the stomatogastric ganglion \cite{gutierrez2013multiple}: a neural circuit whose parametric degeneracy is closely studied \cite{goldman2001global}.
Then, we show EPI's scalability to high dimensional parameter distributions by inferring connectivities of recurrent neural networks (RNNs) that exhibit stable, yet amplified responses -- a hallmark of neural responses throughout the brain \cite{murphy2009balanced, hennequin2014optimal, bondanelli2019population}.}
\q{intro_results2}{In a model of primary visual cortex \cite{litwin2016inhibitory, palmigiano2020structure}, EPI reveals how the recurrent processing across different neuron-type populations shapes excitatory variability: a finding that we show is analytically intractable.
Finally, we investigated the possible connectivities of superior colliculus that allow execution of different tasks on interleaved trials \cite{duan2018collicular}.
EPI discovered a rich distribution containing two connectivity regimes with different solution classes.
We queried the deep probability distribution learned by EPI to produce a mechanistic understanding of neural responses in each regime. 
Intriguingly, the inferred connectivities of each regime reproduced results from optogenetic inactivation experiments in markedly different ways.}
These theoretical insights afforded by EPI illustrate the value of deep inference for the interrogation of neural circuit models.

@@

\q{results_hess}{The structure of the inferred parameter distributions of EPI can be analyzed to reveal key information about how the circuit model produces the emergent property.
The modes of $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ indicate parameter choices emblematic of the emergent property (Fig. \ref{fig:STG}E yellow star).
As probability in the EPI distribution decreases, the emergent property deteriorates.
Perturbing $\mathbf{z}$ along a dimension in which $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ change little will not disturb the emergent property, making this parameter combination \emph{robust} with respect to the emergent property.
In contrast, if $\mathbf{z}$ is perturbed along a dimension with strongly decreasing $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$, that parameter combination is deemed \emph{sensitive} \cite{raue2009structural, raman2017delineating}.
By querying the second order derivative (Hessian) of $\log q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ at a mode, we can quantitatively identify how sensitive (or robust) each eigenvector is by its eigenvalue; the more negative, the more sensitive and the closer to zero, the more robust (see Section \ref{methods_stg_hessian}).
Indeed, samples equidistant from the mode along these dimensions of sensitivity ($\mathbf{v}_1$, smaller eigenvalue) and robustness ($\mathbf{v}_2$, greater eigenvalue) (Fig. \ref{fig:STG}E, arrows) agree with error contours (Fig. \ref{fig:STG}E contours) and have diminished or preserved hub frequency, respectively (Fig. \ref{fig:STG}F activity traces).
The directionality of $\mathbf{v}_2$ suggests that changes in conductance along this parameter combination will most preserve hub neuron firing between the intrinsic rates of the pyloric and gastric mill rhythms.
Importantly, once an EPI distribution has been learned, the modes and Hessians of the distribution can be measured with trivial computation (see Section \ref{methods_NF}).}

In the following sections, we demonstrate EPI on three neural circuit models across ranges of biological realism, neural system function, and network scale.
First, we demonstrate the superior scalability of EPI compared to alternative techniques by inferring high-dimensional distributions of recurrent neural network connectivities that exhibit amplified, yet stable responses.
Next, in a model of primary visual cortex \cite{litwin2016inhibitory, palmigiano2020structure}, we show how EPI discovers parametric degeneracy, revealing how input variability across neuron types affects the excitatory population.
Finally, in a model of superior colliculus \cite{duan2018collicular}, we used EPI to capture multiple parametric regimes of task switching, and queried the dimensions of parameter sensitivity to characterize each regime.

@@

For comparison, we infer the parameters $\mathbf{z}$ likely to produce stable amplification using two alternative simulation-based inference approaches.
Sequential Monte Carlo approximate Bayesian computation (SMC-ABC) \cite{sisson2007sequential} is a rejection sampling approach that uses SMC techniques to improve efficiency, and sequential neural posterior estimation (SNPE) \cite{gonccalves2019training} approximates posteriors with deep probability distributions using a two-network architecture (see Section \ref{methods_related}).
Unlike EPI, these statistical inference techniques do not constrain the statistics of the predictive distribution, so they were run by conditioning on an exemplary dataset  $\mathbf{x}_0 = \bm{\mu}$, following standard practice with these methods \cite{sisson2007sequential, gonccalves2019training}.
To compare the efficiency of these different techniques, we measured the time and number of simulations necessary for the distance of the predictive mean to be less than 0.5 from $\bm{\mu} = \mathbf{x}_0$ (see Section \ref{methods_RNN}).

@@

As the number of neurons $N$ in the RNN, and thus the dimension of the parameter space $\mathbf{z} \in [-1, 1]^{4N}$, is scaled, we see that EPI converges at greater speed and at greater dimension than SMC-ABC and SNPE (Fig. \ref{fig:RNN}A).
It also becomes most efficient to use EPI in terms of simulation count at $N=50$ (Fig. \ref{fig:RNN}B).
It is well known that ABC techniques struggle in parameter spaces of modest dimension \cite{sisson2018handbook}, yet we were careful to assess the scalability of SNPE, which is a more closely related methodology to EPI.
Between EPI and SNPE, we closely controlled the number of parameters in deep probability distributions by dimensionality (Fig. \ref{fig:RNN1}), and tested more aggressive SNPE hyperparameterizations when SNPE failed to converge (Fig. \ref{fig:RNN3}).
In this analysis, we see that deep inference techniques EPI and SNPE are far more amenable to inference of high dimensional RNN connectivities than rejection sampling techniques like SMC-ABC, and that EPI outperforms SNPE in both wall time (elapsed real time) and simulation count.

@@

To understand these differences, note that EPI outperforms SNPE in high dimensions by using gradient information (from $\nabla_\mathbf{z} f(\mathbf{x}; \mathbf{z}) = \nabla_\mathbf{z} [\text{real}(\lambda_1), \lambda_1^s]^\top$).
This choice agrees with recent speculation that such gradient information could improve the efficiency of simulation-based inference techniques \cite{cranmer2020frontier}, as well as reflecting the classic tradeoff between gradient and sampling efficiency (scaling and speed versus generality).
Since gradients of the emergent property statistics are necessary in EPI optimization, gradient tractability is a key criteria when determining the suitability of a simulation-based inference technique.
If the emergent property gradient is efficiently calculated, EPI is a clear choice for inferring high dimensional parameter distributions. 
In the next two sections, we use EPI for novel scientific insight by examining the structure of inferred distributions.

@@

\begin{figure}
\vspace{-1cm}
\begin{center}
\includegraphics[scale=0.8]{figures/fig3/fig3.pdf}
\end{center}
\vspace{-1cm}
\caption{Emergent property inference in the stochastic stabilized supralinear network (SSSN)
\textbf{A}.  Four-population model of primary visual cortex with excitatory (black), parvalbumin (blue), somatostatin (red), and VIP (green) neurons (excitatory and inhibitory projections filled and unfilled, respectively).   
Some neuron-types largely do not form synaptic projections to others ($|W_{\alpha_1, \alpha_2})| < 0.025$).
Each neural population receives a baseline input $\mathbf{h}_b$, and the E- and P-populations also receive a contrast-dependent input $\mathbf{h}_c$.
Additionally, each neural population receives a slow noisy input $\bm{\epsilon}$.
\textbf{B}. Transient network responses of the SSSN model. Traces are independent trials with varying initialization $\mathbf{x}(0)$ and noise $\bm{\epsilon}$. 
\textbf{C}. Mean (solid line) and standard deviation $s_E(\mathbf{x}; \mathbf{z})$ (shading)  across 100 trials.
\textbf{D}. EPI distribution of noise parameters $\mathbf{z}$ conditioned on E-population variability.
The EPI predictive distribution of $s_E(\mathbf{x}; \mathbf{z})$ is show on the bottom-left.
\textbf{E}. (Top) Enlarged visualization of the $\sigma_E$-$\sigma_P$ marginal distribution of EPI $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{ Hz})$ and $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(10\text{ Hz})$.
Each black dot shows the mode at each $\sigma_P$.
The arrows show the most sensitive dimensions of the Hessian evaluated at these modes.
\textbf{F}. The predictive distributions of $\sigma_E^2 + \sigma_P^2$ of each inferred distribution $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{ Hz})$ and $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(10\text{ Hz})$.
}
 \label{fig:V1}
\end{figure}

@@

Figure \ref{fig:V1}E suggests a quadratic relationship in E-population fluctuations and the standard deviation of E- and P-population input; as the square of either $\sigma_E$ or $\sigma_P$ increases, the other compensates by decreasing to preserve the level of $s_E(\mathbf{x}; \mathbf{z})$.
This quadratic relationship is preserved at greater level of E-population variability $\mathcal{X}(10\text{ Hz})$ (Fig. \ref{fig:V1}E).
Indeed, the sum of squares of $\sigma_E$ and $\sigma_P$ is larger in $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(10\text{ Hz}))$ than $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{ Hz}))$ (Fig \ref{fig:V1}F, $p < 1 \times 10^{-10}$), while the sum of squares of $\sigma_S$ and $\sigma_V$ are not significantly different in the two EPI distributions (Fig. \ref{fig:V1_3}, $p=.40$), in which parameters were bounded from 0 to 0.5.
The strong interaction between E- and P-population input variability on excitatory variability is intriguing, since this circuit exhibits a paradoxical effect in the P-population (and no other inhibitory types) (Fig. \ref{fig:V1_3}), meaning that the E-population is P-stabilized.
Future research may uncover a link between the population of network stabilization and compensatory interactions governing excitatory variability.

@@

%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{EPI identifies two regimes of rapid task switching} \label{results_SC}
It has been shown that rats can learn to switch from one behavioral task to the next on randomly interleaved trials \cite{duan2015requirement}, and an important question is what types of neural connectivity produce this computation.
In this experimental setup, rats were explicitly cued on each trial to either orient towards a visual stimulus in the Pro (P) task or orient away from the stimulus in the Anti (A) task (Fig. \ref{fig:SC}A). 
Neural recordings in superior colliculus (SC) exhibited two populations of neurons that represented task context (Pro or Anti).
Furthermore, Pro/Anti neurons in each hemisphere were strongly correlated with the animal's decision \cite{duan2018collicular}.
These results motivated a model of SC that is a four-population dynamical system with functionally-defined neuron-types.
Here, our goal is to understand how connectivity in this circuit model governs the ability to perform rapid task switching: to respond with satisfactory accuracy in both tasks on randomly interleaved trials.

In this SC model, there are Pro- and Anti-populations in each hemisphere (left (L) and right (R)) with activity variables $\mathbf{x} = [x_{LP}, x_{LA}, x_{RP}, x_{RA}]^\top$.
The connectivity of these populations is parameterized by self $sW$, vertical $vW$, diagonal $dW$ and horizontal $hW$ connections (Fig. \ref{fig:SC}B).
The input $\mathbf{h}$ is comprised of a positive cue-dependent signal to the Pro or Anti populations, a positive stimulus-dependent input to either the Left or Right populations, and a choice-period input to the entire network (see Section \ref{methods_sc_model}).
Model responses are bounded from 0 to 1 as a function $\phi$ of an internal variable $\mathbf{u}$
\begin{equation}
\begin{split}
\tau \frac{d\mathbf{u}}{dt} &= -\mathbf{u} + W\mathbf{x} + \mathbf{h} + d\mathbf{B} \\
\mathbf{x} &= \phi(\mathbf{u}).
\end{split}
\end{equation}
The model responds to the side with greater Pro neuron activation; e.g. the response is left if $x_{LP} > x_{RP}$ at the end of the trial.  
Here, we use EPI to determine the network connectivity $\mathbf{z} = [sW, vW, dW, hW]^{\top}$ that produces rapid task switching.

\begin{figure}
\begin{center}
\includegraphics[scale=0.8]{figures/fig4/fig4.pdf}
\end{center}
\caption{ 
\textbf{A}. Rapid task switching behavioral paradigm (see text). 
\textbf{B}. Model of superior colliculus (SC). Neurons: LP - Left Pro, RP - Right Pro, LA - Left Anti, RA - Right Anti. 
Parameters: $sW$ - self, $hW$ - horizontal, $vW$ -vertical, $dW$ - diagonal weights.  
\textbf{C}. The EPI inferred distribution of rapid task switching networks.  
Stars indicate modes $\mathbf{z}^*$ whose color indicates connectivity regime (see legend Fig \ref{fig:SC}D).
Sensitivity vectors $\mathbf{v}_1(\mathbf{z}^*)$ are shown by arrows.
(Bottom-left) EPI predictive distribution of task accuracies.
\textbf{D}. The connectivity regimes have different responses to perturbation.
(Top) Mean and standard error ($N_{\text{test}}$ = 25) of accuracy with respect to perturbation along the sensitivity dimension of each mode $\mathbf{z}^*$.
(Middle) Same with perturbation in the dimension of increasing $\lambda_{\text{task}}$ ($\mathbf{v}_{\text{task}}$).
(Bottom) Same with perturbation in the dimension of increasing $\lambda_{\text{diag}}$ ($\mathbf{v}_{\text{diag}}$).
}
\label{fig:SC}
\end{figure}

@@

The EPI inferred distribution  (Fig. \ref{fig:SC}C) produces Pro and Anti task accuracies (Fig. \ref{fig:SC}C, middle-left) consistent with rapid task switching (Equation \ref{eq:SC_EP}).
This parameter distribution has rich structure, that is not captured well by simple linear correlations (Fig. \ref{fig:SC1}).
Specifically, the shape of the EPI distribution is sharply bent, matching ground truth structure indicated by brute-force sampling (Fig. \ref{fig:SC8})
This is most saliently pointed out in the marginal distribution of  $sW$-$hW$ (Fig. \ref{fig:SC}C top-right), where anticorrelation between $sW$ and $hW$ switches to correlation with decreasing $sW$.
The two regimes produce different types of responses in the Pro and Anti tasks (Fig. SC2).
Not only has EPI captured this complicated distribution of connectivities producing rapid task switching, we can query the EPI distribution $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ to understand these two parametric regimes of SC connectivity.

To distinguish these two regimes, we use the EPI distribution to identify two sets of modes.
By fixing $hW$ to different values and doing gradient ascent on $\log q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$, we arrive at two solutions $\mathbf{z}^*(hW_{\text{fixed}}, r)$ where regime $r \in [1,2]$, and regime 1 is that of greater $sW$ (see Section \ref{methods_sc_regime}).
As $hW_{\text{fixed}}$ increases, the modes coalesce to intermediate parameters reflecting a transition between the two sets of modes (Fig. \ref{fig:SC4} top).
By using EPI to connect these two regimes through this transitionary region of parameter space, we can explore what distinguishes the two regimes by stepping from the prototypical connectivity of regime 1 to that of regime 2.

Although the connectivities gradually coalesce to the transitionary part of parameter space, the sensitivity dimensions $\mathbf{v}_1(\mathbf{z})$ are categorically different across regimes (Fig. \ref{fig:SC4} bottom).
The sensitivity dimension identifies the parameter combination which causes the emergent property to diminish with the shortest perturbation.
Since the two regimes have different $\mathbf{v}_1(\mathbf{z})$, this suggests they have different pathologies in their connectivity.
By perturbing connectivity in each regime along the sensitivity dimension, we can get a sense of the differing nature of these pathologies.

When perturbing connectivity along the sensitivity dimension, Pro accuracy monotonically increases in both regimes (Fig. \ref{fig:SC}D, top-left).
However, there is a stark difference between regimes in Anti accuracy. 
Anti accuracy falls in either direction of $\mathbf{v}_1$ in regime 1, yet monotonically increases along with Pro accuracy in regime 2 (Fig. \ref{fig:SC}D, top-right).
These distinct pathologies of rapid task switching are caused by distinct connectivity changes ($\mathbf{v}_1(\mathbf{z}^*(r=1))$ vs $\mathbf{v}_1(\mathbf{z}^*(r=2))$) and explain the sharp change in local structure of the EPI distribution.

>>>>

\newcommand\startsupplement{%
    \makeatletter 
       \setcounter{table}{0}
       \renewcommand{\thetable}{S\arabic\c@table}
       \setcounter{figure}{0}
       \renewcommand{\thefigure}{S\arabic{figure}}
    \makeatother}

@@

Equipped with this method, we demonstrate the capabilities of EPI and present novel theoretical findings from its analysis.
\q{intro_results1}{First, we show EPI's ability to handle biologically realistic circuit models using a five-neuron model of the stomatogastric ganglion \cite{gutierrez2013multiple}: a neural circuit whose parametric degeneracy is closely studied \cite{goldman2001global}.
Then, we show EPI's scalability to high dimensional parameter distributions by inferring connectivities of recurrent neural networks (RNNs) that exhibit stable, yet amplified responses -- a hallmark of neural responses throughout the brain \cite{murphy2009balanced, hennequin2014optimal, bondanelli2019population}.}
\q{intro_results2}{In a model of primary visual cortex \cite{litwin2016inhibitory, palmigiano2020structure}, EPI reveals how the recurrent processing across different neuron-type populations shapes excitatory variability: a finding that we show is analytically intractable.
Finally, we investigated the possible connectivities of superior colliculus that allow execution of different tasks on interleaved trials \cite{duan2019collicular}.
EPI discovered a rich distribution containing two connectivity regimes with different solution classes.
We queried the deep probability distribution learned by EPI to produce a mechanistic understanding of neural responses in each regime. 
Intriguingly, the inferred connectivities of each regime reproduced results from optogenetic inactivation experiments in markedly different ways.}
These theoretical insights afforded by EPI illustrate the value of deep inference for the interrogation of neural circuit models.

@@

\q{results_hess}{The structure of the inferred parameter distributions of EPI can be analyzed to reveal key information about how the circuit model produces the emergent property.
As probability in the EPI distribution decreases away from the mode of $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ (Fig. \ref{fig:STG}E yellow star), the emergent property deteriorates.
Perturbing $\mathbf{z}$ along a dimension in which $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ change little will not disturb the emergent property, making this parameter combination \emph{robust} with respect to the emergent property.
In contrast, if $\mathbf{z}$ is perturbed along a dimension with strongly decreasing $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$, that parameter combination is deemed \emph{sensitive} \cite{raue2009structural, raman2017delineating}.
By querying the second order derivative (Hessian) of $\log q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X})$ at a mode, we can quantitatively identify how sensitive (or robust) each eigenvector is by its eigenvalue; the more negative, the more sensitive and the closer to zero, the more robust (see Section \ref{methods_stg_hessian}).
Indeed, samples equidistant from the mode along these dimensions of sensitivity ($\mathbf{v}_1$, smaller eigenvalue) and robustness ($\mathbf{v}_2$, greater eigenvalue) (Fig. \ref{fig:STG}E, arrows) agree with error contours (Fig. \ref{fig:STG}E contours) and have diminished or preserved hub frequency, respectively (Fig. \ref{fig:STG}F activity traces).
The directionality of $\mathbf{v}_2$ suggests that changes in conductance along this parameter combination will most preserve hub neuron firing between the intrinsic rates of the pyloric and gastric mill rhythms.
Importantly, once an EPI distribution has been learned, the modes and Hessians of the distribution can be measured with trivial computation (see Section \ref{methods_NF}).}

In the following sections, we demonstrate EPI on three neural circuit models across ranges of biological realism, neural system function, and network scale.
First, we demonstrate the superior scalability of EPI compared to alternative techniques by inferring high-dimensional distributions of recurrent neural network connectivities that exhibit amplified, yet stable responses.
Next, in a model of primary visual cortex \cite{litwin2016inhibitory, palmigiano2020structure}, we show how EPI discovers parametric degeneracy, revealing how input variability across neuron types affects the excitatory population.
Finally, in a model of superior colliculus \cite{duan2019collicular}, we used EPI to capture multiple parametric regimes of task switching, and queried the dimensions of parameter sensitivity to characterize each regime.

@@

For comparison, we infer the parameters $\mathbf{z}$ likely to produce stable amplification using two alternative simulation-based inference approaches.
Sequential Monte Carlo approximate bayesian computation (SMC-ABC) \cite{sisson2007sequential} is a rejection sampling approach that uses SMC techniques to improve efficiency, and sequential neural posterior estimation (SNPE) \cite{gonccalves2019training} approximates posteriors with deep probability distributions using a two-network architecture (see Section \ref{methods_related}).
Unlike EPI, these statistical inference techniques do not constrain the statistics of the predictive distribution, so they were run by conditioning on an exemplary dataset  $\mathbf{x}_0 = \bm{\mu}$, following standard practice with these methods \cite{sisson2007sequential, gonccalves2019training}.
To compare the efficiency of these different techniques, we measured the time and number of simulations necessary for the distance of the predictive mean to be less than 0.5 from $\bm{\mu} = \mathbf{x}_0$ (see Section \ref{methods_RNN}).

@@

As the number of neurons $N$ in the RNN, and thus the dimension of the parameter space $\mathbf{z} \in [-1, 1]^{4N}$, is scaled, we see that EPI converges at greater speed and at greater dimension than SMC-ABC and SNPE (Fig. \ref{fig:RNN}A).
It also becomes most efficient to use EPI in terms of simulation count at $N=50$ (Fig. \ref{fig:RNN}B).
It is well known that ABC techniques struggle in parameter spaces of modest dimension \cite{sisson2018handbook}, yet we were careful to assess the scalability of SNPE, which is a more closely related methodology to EPI.
Between EPI and SNPE, we closely controlled the number of parameters in deep probability distributions by dimensionality (Fig. \ref{fig:RNN1}), and tested more aggressive SNPE hyperparameter choices when SNPE failed to converge (Fig. \ref{fig:RNN3}).
In this analysis, we see that deep inference techniques EPI and SNPE are far more amenable to inference of high dimensional RNN connectivities than rejection sampling techniques like SMC-ABC, and that EPI outperforms SNPE in both wall time (elapsed real time) and simulation count.

@@

To understand these differences, note that EPI outperforms SNPE in high dimensions by using gradient information (from $\nabla_\mathbf{z} f(\mathbf{x}; \mathbf{z}) = \nabla_\mathbf{z} [\text{real}(\lambda_1), \lambda_1^s]^\top$).
This choice agrees with recent speculation that such gradient information could improve the efficiency of simulation-based inference techniques \cite{cranmer2020frontier}, as well as reflecting the classic tradeoff between gradient-based and sampling-based estimators (scaling and speed versus generality).
Since gradients of the emergent property statistics are necessary in EPI optimization, gradient tractability is a key criteria when determining the suitability of a simulation-based inference technique.
If the emergent property gradient is efficiently calculated, EPI is a clear choice for inferring high dimensional parameter distributions. 
In the next two sections, we use EPI for novel scientific insight by examining the structure of inferred distributions.

@@

\begin{figure}
\vspace{-1cm}
\begin{center}
\includegraphics[scale=0.8]{figures/fig3/fig3.pdf}
\end{center}
\caption{Emergent property inference in the stochastic stabilized supralinear network (SSSN)
\textbf{A}.  Four-population model of primary visual cortex with excitatory (black), parvalbumin (blue), somatostatin (red), and VIP (green) neurons (excitatory and inhibitory projections filled and unfilled, respectively).   
Some neuron-types largely do not form synaptic projections to others ($|W_{\alpha_1, \alpha_2})| < 0.025$).
Each neural population receives a baseline input $\mathbf{h}_b$, and the E- and P-populations also receive a contrast-dependent input $\mathbf{h}_c$.
Additionally, each neural population receives a slow noisy input $\bm{\epsilon}$.
\textbf{B}. Transient network responses of the SSSN model. Traces are independent trials with varying initialization $\mathbf{x}(0)$ and noise $\bm{\epsilon}$. 
\textbf{C}. Mean (solid line) and standard deviation $s_E(\mathbf{x}; \mathbf{z})$ (shading)  across 100 trials.
\textbf{D}. EPI distribution of noise parameters $\mathbf{z}$ conditioned on E-population variability.
The EPI predictive distribution of $s_E(\mathbf{x}; \mathbf{z})$ is show on the bottom-left.
\textbf{E}. (Top) Enlarged visualization of the $\sigma_E$-$\sigma_P$ marginal distribution of EPI $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{ Hz})$ and $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(10\text{ Hz})$.
Each black dot shows the mode at each $\sigma_P$.
The arrows show the most sensitive dimensions of the Hessian evaluated at these modes.
\textbf{F}. The predictive distributions of $\sigma_E^2 + \sigma_P^2$ of each inferred distribution $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{ Hz})$ and $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(10\text{ Hz})$.
}
 \label{fig:V1}
\end{figure}

@@

Figure \ref{fig:V1}E suggests a quadratic relationship in E-population fluctuations and the standard deviation of E- and P-population input; as the square of either $\sigma_E$ or $\sigma_P$ increases, the other compensates by decreasing to preserve the level of $s_E(\mathbf{x}; \mathbf{z})$.
This quadratic relationship is preserved at greater level of E-population variability $\mathcal{X}(10\text{ Hz})$ (Fig. \ref{fig:V1}E and \ref{fig:V1_1}).
Indeed, the sum of squares of $\sigma_E$ and $\sigma_P$ is larger in $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(10\text{ Hz}))$ than $q_{\bm{\theta}}(\mathbf{z} \mid \mathcal{X}(5\text{ Hz}))$ (Fig \ref{fig:V1}F, $p < 1 \times 10^{-10}$), while the sum of squares of $\sigma_S$ and $\sigma_V$ are not significantly different in the two EPI distributions (Fig. \ref{fig:V1_3}, $p=.40$), in which parameters were bounded from 0 to 0.5.
The strong interaction between E- and P-population input variability on excitatory variability is intriguing, since this circuit exhibits a paradoxical effect in the P-population (and no other inhibitory types) (Fig. \ref{fig:V1_4}), meaning that the E-population is P-stabilized.
Future research may uncover a link between the population of network stabilization and compensatory interactions governing excitatory variability.

@@

%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{EPI identifies two regimes of rapid task switching} \label{results_SC}
It has been shown that rats can learn to switch from one behavioral task to the next on randomly interleaved trials \cite{duan2015requirement}, and an important question is what neural mechanisms produce this computation.
In this experimental setup, rats were given an explicit task cue on each trial, either Pro or Anti. 
After a delay period, rats were shown a stimulus, and made a context  (task) dependent response (Fig. \ref{fig:SC}A). 
In the Pro task, rats were required to orient towards the stimulus, while in the Anti task, rats were required to orient away from the stimulus.
Pharmacological inactivation of the SC impaired rat performance, and time-specific optogenetic inactivation revealed a crucial role for the SC on the cognitively demanding Anti trials \cite{duan2019collicular}.
These results motivated a nonlinear dynamical model of the SC containing four functionally-defined neuron-type populations.
In Duan et al. 2019, a computationally intensive procedure was used to obtain a set of 373 connectivity parameters that qualitatively reproduced these optogenetic inactivation results.
To build upon the insights of this previous work, we use the probabilistic tools afforded by EPI to identify and mechanistically characterize two linked, yet distinct regimes of rapid task switching connectivity.

In this SC model, there are Pro- and Anti-populations in each hemisphere (left (L) and right (R)) with activity variables $\mathbf{x} = [x_{LP}, x_{LA}, x_{RP}, x_{RA}]^\top$ \cite{duan2019collicular}.
The connectivity of these populations is parameterized by self $sW$, vertical $vW$, diagonal $dW$ and horizontal $hW$ connections (Fig. \ref{fig:SC}B).
The input $\mathbf{h}$ is comprised of a positive cue-dependent signal to the Pro or Anti populations, a positive stimulus-dependent input to either the Left or Right populations, and a choice-period input to the entire network (see Section \ref{methods_sc_model}).
Model responses are bounded from 0 to 1 as a function $\phi$ of an internal variable $\mathbf{u}$
\begin{equation}
\begin{split}
\tau \frac{d\mathbf{u}}{dt} &= -\mathbf{u} + W\mathbf{x} + \mathbf{h} + d\mathbf{B} \\
\mathbf{x} &= \phi(\mathbf{u}).
\end{split}
\end{equation}
The model responds to the side with greater Pro neuron activation; e.g. the response is left if $x_{LP} > x_{RP}$ at the end of the trial.  
Here, we use EPI to determine the network connectivity $\mathbf{z} = [sW, vW, dW, hW]^{\top}$ that produces rapid task switching.

\begin{figure}
\begin{center}
\includegraphics[scale=0.75]{figures/fig4/fig4.pdf}
\end{center}
\vspace{-.25cm}
\caption{\small
\textbf{A}. Rapid task switching behavioral paradigm (see text). 
\textbf{B}. Model of superior colliculus (SC). Neurons: LP - Left Pro, RP - Right Pro, LA - Left Anti, RA - Right Anti. 
Parameters: $sW$ - self, $hW$ - horizontal, $vW$ -vertical, $dW$ - diagonal weights.  
\textbf{C}. The EPI inferred distribution of rapid task switching networks.  
Red/purple parameters indicate modes $\mathbf{z}^*(sW)$ colored by $sW$.
Sensitivity vectors $\mathbf{v}_1(\mathbf{z}^*)$ are shown by arrows.
(Bottom-left) EPI predictive distribution of task accuracies.
\textbf{D}. Mean and standard error ($N_{\text{test}}$ = 25) of accuracy in Pro (top) and Anti (bottom) tasks after perturbing connectivity away from mode along $\mathbf{v}_1(\mathbf{z}^*)$ (left), $\mathbf{v}_{\text{task}}$ (middle), and $\mathbf{v}_{\text{diag}}$ (right).
}
\label{fig:SC}
\end{figure}

%\begin{figure}
%\floatbox[{\capbeside\thisfloatsetup{capbesideposition={right,top},capbesidewidth=5cm}}]{figure}[\FBwidth]
%{\caption{\small
%\textbf{A}. Rapid task switching behavioral paradigm (see text). 
%\textbf{B}. Model of superior colliculus (SC). Neurons: LP - Left Pro, RP - Right Pro, LA - Left Anti, RA - Right Anti. 
%Parameters: $sW$ - self, $hW$ - horizontal, $vW$ -vertical, $dW$ - diagonal weights.  
%\textbf{C}. The EPI inferred distribution of rapid task switching networks.  
%Red/purple parameters indicate modes $\mathbf{z}^*(sW)$ colored by $sW$.
%Sensitivity vectors $\mathbf{v}_1(\mathbf{z}^*)$ are shown by arrows.
%(Bottom-left) EPI predictive distribution of task accuracies.
%\textbf{D}. Mean and standard error ($N_{\text{test}}$ = 25) of accuracy in Pro (top) and Anti (bottom) tasks after perturbing connectivity away from mode $\mathbf{z}^*(sW)$ by distance $\delta$ along $\mathbf{v}_1(\mathbf{z}^*)$ (left), $\mathbf{v}_{\text{task}}$ (middle), and $\mathbf{v}_{\text{diag}}$ (right).
%}
%\label{fig:test}}
%{\hspace{-.5in}\includegraphics[scale=.73]{figures/fig4/fig4.pdf}}
%\end{figure}

@@

The EPI inferred distribution  (Fig. \ref{fig:SC}C) produces Pro and Anti task accuracies (Fig. \ref{fig:SC}C, bottom-left) consistent with rapid task switching (Equation \ref{eq:SC_EP}).
This parameter distribution has rich structure, that is not captured well by simple linear correlations (Fig. \ref{fig:SC1}).
Specifically, the shape of the EPI distribution is sharply bent, matching ground truth structure indicated by brute-force sampling (Fig. \ref{fig:SC8}).
This is most saliently observed in the marginal distribution of  $sW$-$hW$ (Fig. \ref{fig:SC}C top-right), where anticorrelation between $sW$ and $hW$ switches to correlation with decreasing $sW$.
By identifying the modes of the EPI distribution $\mathbf{z}^*(sW)$ at different values of $sW$ (Fig. \ref{fig:SC}C red/purple dots), we can quantify this change in distributional structure with the sensitivity dimension $\mathbf{v}_1(\mathbf{z})$ (Fig. \ref{fig:SC}C red/purple arrows).
Note that the directionality of these sensitivity dimensions at $\mathbf{z}^*(sW)$ changes distinctly with $sW$, and are perpendicular to the flat dimensions of the EPI distribution that preserve rapid task switching.
These two directionalities of sensitivity motivate the distinction of connectivity into two regimes, which produce different types of responses in the Pro and Anti tasks (Fig. \ref{fig:SC2}).

When perturbing connectivity along the sensitivity dimension away from the modes
\begin{equation}
\mathbf{z} = \mathbf{z}^*(sW) + \delta \mathbf{v}_1(\mathbf{z}^*(sW)),
\end{equation}
Pro accuracy monotonically increases in both regimes (Fig. \ref{fig:SC}D, top-left).
However, there is a stark difference between regimes in Anti accuracy. 
Anti accuracy falls in either direction of $\mathbf{v}_1$ in regime 1, yet monotonically increases along with Pro accuracy in regime 2 (Fig. \ref{fig:SC}D, bottom-left).
The sharp change in local structure of the EPI distribution is therefore explained by distinct sensitivities: Anti accuracy diminishes in only one or both directions of the sensitivity perturbation.
